---
title: "Mini Data Analysis: Report 1"
author: "Harsh Sharma"
date: "09/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This report includes the exploratory data analysis performed on a dataset chosen from the `datateacher` package. Incremental effort will be placed in familiarizing with different data sets to finally choose one, and four research questions will formulated around it.

## Set-up
First we need to load the following two packages:

```{r, message=FALSE}
library(datateachr)
library(tidyverse)
```

The `datateachr` is comprised of the 7 semi-tidy datasets:

> 1. apt_buildings: Acquired courtesy of The City of Toronto’s Open Data Portal. It currently has 3455 rows and 37 columns.

> 2. building_permits: Acquired courtesy of The City of Vancouver’s Open Data Portal. It currently has 20680 rows and 14 columns.

> 3. cancer_sample: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

> 4. flow_sample: Acquired courtesy of The Government of Canada’s Historical Hydrometric Database. It currently has 218 rows and 7 columns.

> 5. parking_meters: Acquired courtesy of The City of Vancouver’s Open Data Portal. It currently has 10032 rows and 22 columns.

> 6. steam_games: Acquire d courtesy of Kaggle. It currently has 40833 rows and 21 columns.

> 7. vancouver_trees: Acquired courtesy of The City of Vancouver’s Open Data Portal. It currently has 146611 rows and 20 columns.

## Task 1.1: Priliminary Dataset Selection

Based on the descriptions, I narrow down my intial choices to the following:

1. *building_permits:* working for a regulatory agency, I have prior experience working in compliance of permits and licenses.

2. *cancer_sample:* I am interested to pursue my research in a field related to algorithms employed in healthcare digital images so this seems very meaningful and applicable

3. *parking_meters:* having worked for Translink, this may be an interesting dataset to probe into parking trends and by extension into travel behavior

4. *vancouver_trees:* interesting dataset that may be leverage-able to assess the effects of climate change and/or rate of planting by the city

## Task 1.3: Intermediate Dataset Selection

Based on the above I further limit my choices to the following top contenders:

1.  cancer_sample

2. 

## Task 1.4: Final Dataset Selection

In order to make the concluding dataset choice, it is important to consider the end result i.e. the associated research question. This will help gauge the project that is most appealing. Following are my research questions:

1. cancer_sample