---
title: "Mini Data Analysis: Report 1"
author: "Harsh Sharma"
date: "09/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This report includes the exploratory data analysis performed on a dataset chosen from the `datateacher` package. Incremental effort will be placed in familiarizing with different data sets to finally choose one, and four research questions will formulated around it.

## Set-up
First we need to load the following two packages:

```{r, message=FALSE}
library(datateachr)
library(tidyverse)
```

The `datateachr` is comprised of the 7 semi-tidy datasets:

> 1. apt_buildings: Acquired courtesy of The City of Toronto’s Open Data Portal. It currently has 3455 rows and 37 columns.

> 2. building_permits: Acquired courtesy of The City of Vancouver’s Open Data Portal. It currently has 20680 rows and 14 columns.

> 3. cancer_sample: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

> 4. flow_sample: Acquired courtesy of The Government of Canada’s Historical Hydrometric Database. It currently has 218 rows and 7 columns.

> 5. parking_meters: Acquired courtesy of The City of Vancouver’s Open Data Portal. It currently has 10032 rows and 22 columns.

> 6. steam_games: Acquire d courtesy of Kaggle. It currently has 40833 rows and 21 columns.

> 7. vancouver_trees: Acquired courtesy of The City of Vancouver’s Open Data Portal. It currently has 146611 rows and 20 columns.

## Task 1: Dataset Selection

**1.1 Based on the descriptions, I narrow down my intial choices to the following:**

1. building_permits:
Working for a regulatory agency, I have prior experience working in compliance of permits and licenses
2. 