---
title: "Mini Data Analysis: Report 2"
author: "Harsh Sharma"
date: "21/10/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This report continues the exploratory data analysis performed in Report1 on the 'cancer_dataset' chosen from the `datateacher` package. Incremental effort will be placed in exploring in more detail the dataset in a fashion that will lead to geenrating insights to answer the research questions.

## Set-up
First we need to load the following two packages:

```{r, message=FALSE}
# Installing packages if missing, required only once
# install.packages("devtools")
# devtools::install_github("UBC-MDS/datateachr")

library(datateachr)
library(tidyverse)
```

## Task1: Processing and Summarizing Data

```{r, include=FALSE, message=FALSE}
# Supporting in-line code below
dim <- dim(cancer_sample)
class <- class(cancer_sample)
```

From earlier analysis we know the cancer_dataset has the following high level characteristics:

  - Shape is: `r dim`
  
  - Data class is: `r class`
  
### Task 1.1: Research Questions

Following are the research questions formulated in Report1:

1. **Can a binary (malignant or benign) classification model be generated the given variables and dataset?**

2.  **Is data distribution/characteristics comparable between the benign and malignant classes?**

3.  **What variable(s) have the strongest correlations to the response variable?**
Which next leads us to finding the optimized set of variables having the highest model performance. This inherently involves analyzing the three associated columns for the same parameter (example: radius\_mean, radius\_se, radius\_worst) in order to minimize redundancy.

4.  **Identifying what variable(s) contributes to the highest model sensitivity**.
In this problem setting it is better to have a false positive than a false negative.

**PS:** Research question 4 has been replaced in the follwong section due to better alignment with this report's instruction and also the course scope.

### Task 1.2: In-depth EDA
In this section atleast one summarizing and one graphing task is attempted per research question.

#### Question1 - Generating binary classification model
Computing the summary statistic of numeric variable 'radius_mean' across the categorical variable 'diagnosis':

```{r}
# We need to first group by diagnosis
cancer_sample %>%
  group_by(diagnosis) %>%
  summarize(min_value=min(radius_mean), max_value=max(radius_mean), mean=mean(radius_mean,na.rm=TRUE),median=median(radius_mean), sd=sd(radius_mean))
```


### Task1.3: Learning